From Freebase to Wikidata: The Great Migration

### 2015-10-15 paper.tex
Change Thomas's email address
Minor corrections in affiliations and style/grammar.

\numberofauthors{5}
\author{
\alignauthor
Thomas Pellissier Tanon\titlenote{The author was an intern at Google when the majority of the work was done.}
   \affaddr{Google, San Francisco, USA}\\   
   \email{thomaspelta@gmail.com}
\alignauthor
Denny Vrandečić
   \affaddr{Google, San Francisco, USA}\\   
   \email{vrandecic@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google, Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\and
\alignauthor
Thomas Steiner\\
   \affaddr{Google, Hamburg, Germany}\\   
   \email{tomac@google.com}
\alignauthor
Lydia Pintscher\\
   \affaddr{Wikimedia, Berlin, Germany}\\   
   \email{mail@pintscher.de}
}
\begin{document}

@@

In order to prepare the data for integration into the Primary Sources Tool,
we have created a set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code is available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping tool.
We first created a~small dataset of around 1M statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.
In the following, we progressively added more statements to the tool's database back-end.
Thereby, we were able to slowly upscale the tool's database back-end
without risking back-end performance bottlenecks.
The process of adding data in small batches of 1M~to~2M statements per batch
further allowed us to detect some potential issues in the front-end
and fix them before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts easily
and correct minor issues with the data early on already in the back-end.
Upon approval or rejection of either a~claim or a~reference,
the Wikidata item page in question reloads in order to reflect
the potentially changed state of the migration, as,
in case of incoming claims with references, freshly accepted claims 
can require incoming references to be attached elsewhere in the hierarchy.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\columnwidth]{img/architecture.png}
    \caption{Primary Sources Tool architecture. The front-end is implemented in
      JavaScript and sends HTTP requests to the C++ back-end to retrieve and
      update entities or statements, which are stored in a relational database.
      \todo{Create PDF version of the diagram, improve consistency of used terms}}
    \label{fig:architecture}
\end{figure}

Entities and statements are stored in a relational database (MySQL or SQLite)
and follow the Wikidata data structure with an additional status field
to keep track of statements that either already have been approved or rejected,
or not yet visited.
Statements can be grouped in ``datasets'' and ``uploads'' to
distinguish data from different sources.\footnote{At the time of this writing,
the back-end contained five different datasets.}

@@

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.
The used method is to extract the earliest date and the longitude
linked to each of the entities of the two knowledge bases
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution \todo{Please clarify the ``propagate'' part}.
We see that the two knowledge bases follow the same basic patterns,
with their coverage mostly concentrated on Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems better in Freebase.
Our theory is that this is \emph{(i)}~due to the absence of a~lot of musical data in Wikidata
that in contrast is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and more even in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.
One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly note that for knowledge bases the amount of triples is not an adequate measure
to judge the proportion of additional good data.

>>>>

\numberofauthors{5}
\author{
\alignauthor
Thomas Pellissier Tanon\titlenote{The author was an intern at Google when the majority of the work was done.}
   \affaddr{Google, San Francisco, USA}\\   
   \email{\mbox{thomas@pellissier-tanon.fr}}
\alignauthor
Denny Vrandečić\\
   \affaddr{Google, San Francisco, USA}\\   
   \email{vrandecic@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google, Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\and
\alignauthor
Thomas Steiner\\
   \affaddr{Google, Hamburg, Germany}\\   
   \email{tomac@google.com}
\alignauthor
Lydia Pintscher\\
   \affaddr{Wikimedia, Berlin, Germany}\\   
   \email{mail@pintscher.de}
}
\begin{document}

@@

In order to prepare the data for integration into the Primary Sources Tool,
we have created a~set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code is available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping tool.
We first created a~small dataset of around 1M statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.
In the following, we progressively added more statements to the tool's database back-end.
Thereby, we were able to slowly upscale the tool's database back-end
without risking back-end performance bottlenecks.
The process of adding data in small batches of 1M~to~2M statements per batch
further allowed us to detect some potential issues in the front-end
and fix them before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts easily
and correct minor issues with the data early on already in the back-end.
Upon approval or rejection of either a~claim or a~reference,
the Wikidata item page in question reloads in order to reflect
the potentially changed state of the migration, as,
in case of incoming claims with references, freshly accepted claims 
can require incoming references to be attached elsewhere in the hierarchy.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\columnwidth]{img/architecture.png}
    \caption{Primary Sources Tool architecture. The front-end is implemented in
      JavaScript and sends HTTP requests to the C++ back-end to retrieve and
      update entities or statements, which are stored in a~relational database.
      \todo{Create PDF version of the diagram, improve consistency of used terms}}
    \label{fig:architecture}
\end{figure}

Entities and statements are stored in a~relational database (MySQL or SQLite)
and follow the Wikidata data structure with an additional status field
to keep track of statements that either already have been approved or rejected,
or not yet visited.
Statements can be grouped in ``datasets'' and ``uploads'' to
distinguish data from different sources.\footnote{At the time of this writing,
the back-end contained five different datasets.}

@@

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.
The used method is to extract the earliest date and the longitude
linked to each of the entities of the two knowledge bases
and propagate them to the connected entities that do not have such data.
To implement this propagation, we set as value (date or longitude)
the average of the values of the entities whose distance is $<10$
with $2^{-d}$ as weighting, $d$~being the distance between the two entities.
We see that the two knowledge bases follow the same basic patterns,
with their coverage mostly concentrated on Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems better in Freebase.
Our theory is that this is \emph{(i)}~due to the absence of a~lot of musical data in Wikidata
that in contrast is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and more even in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.
One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly note that for knowledge bases the amount of triples is not an adequate measure
to judge the proportion of additional good data.
